---
title: 'Capstone: Bellabeat Case Study'
author: "Upasana Sharma"
date: "3/16/2022"
output:
  html_document: default
  pdf_document: default
---

## Introduction: 

This is a case study of a company 'Bellabeat' which is a high-tech manufacturer of health focused products for women. The stakeholders of the company want to  analyze smart device fitness data could help unlock new growth opportunities for the company. 

## Product - Time (Bellabeat Smart Watch): 

This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

## Business Task: 
- To recommend the marketing strategy/high-level recommendation for the one of the Bellabeat smart device (smart watch in this case) based on the analysis of Fitbit users' data. 
- Analyse trends of another smart device data to gain insights on non-bellabeat users.

## Stakeholders:
- Urska Srsen: Bellabeat's cofounder and Chief Creative officer.
- Sando Mur: Bellabeat's cofounder
- Marketing Analytics Team

## Ask:
- What are some trends in smart device usage?
- How could these trends apply to Bellabeat customers?
- How could these trends help influence Bellabeat marketing strategy?

### Database used:
Kaggle public dataset- Fitbit Fitness Tracker Data
(This database has 18 datasets, I have used 3 of them daily_activity, sleep_day and weight_log datasets for my analysis.)
 
### Data Limitations and biases:
- The dataset has only 33 fitbit users and the gender ratio is unknown.
- This data is an year old at time time of this analysis, therefore the trends observed might vary from the current ones. 
- No data available for a particular smart device trends.

## Preparing the data:
Used R language and Rstudio for this analysis.

Loading all necessary libraries
```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```

Reading CSV files into variables
```{r}
daily_activity <- read.csv("C:/Data/dailyActivity_merged.csv")
sleep_day <- read.csv("C:/Data/sleepDay_merged.csv")
weightlog_info <- read.csv("C:/Data/weightLogInfo_merged.csv")
```

Exploring Tables

Table - dailyActivity_merged
```{r}
head(daily_activity)
str(daily_activity)
```

Table - sleepDay
```{r}
head(sleep_day)
str(sleep_day)
```

Table - weightInfo
```{r}
head(weightlog_info)
str(weightlog_info)
```

Understanding statistics summary of all three tables
```{r}
n_distinct(daily_activity$Id)
n_distinct(sleep_day$Id)
n_distinct(weightlog_info$Id)
```

## Processing the data -

Changing the date format - daily_activity
```{r}
daily_activity$ActivityDate = mdy(daily_activity$ActivityDate)
```

Adding a new column 'day_name' for the corresponding dates 
```{r}
daily_act <- daily_activity %>%
  mutate(day_name = wday(daily_activity$ActivityDate, label = TRUE))
```

Creating a new daily activity table summarized by days of the week
```{r}
days_daily_activity <- daily_act %>%
  group_by(day_name) %>%
  summarise(total_steps = sum(TotalSteps), total_distance = sum(TotalDistance), total_calories = sum(Calories), Very_active_dist = sum(VeryActiveDistance),
            mod_active_dist = sum(ModeratelyActiveDistance), light_active_dist = sum(LightActiveDistance), sed_active_dist = sum(SedentaryActiveDistance), 
            very_active_min = sum(VeryActiveMinutes), fairly_active_min = sum(FairlyActiveMinutes), lightly_active_mins = sum(LightlyActiveMinutes),
            sed_min = sum(SedentaryMinutes)) %>%
  select(day_name, total_steps, total_distance, total_calories, Very_active_dist, mod_active_dist, light_active_dist, sed_active_dist, very_active_min,
         fairly_active_min, lightly_active_mins, sed_min)
```

View the created table
```{r}
print(days_daily_activity)
```

Changing the format of date - sleep_Day
```{r}
sleep_day$SleepDay = mdy_hms(sleep_day$SleepDay)
```

Adding a new column 'day_name' for the corresponding dates
```{r}
sleep_d <- sleep_day %>%
  mutate(day_name = wday(SleepDay, label = TRUE))
```

Creating a new sleep_day table summarized by days of the week
```{r}
days_sleep_day <- sleep_d %>%
  group_by(day_name) %>%
  summarise(tot_sleep_rec = sum(TotalSleepRecords), tot_min_asleep = sum(TotalMinutesAsleep), tot_time_bed = sum(TotalTimeInBed)) %>%
  select(day_name, tot_sleep_rec, tot_min_asleep, tot_time_bed)
```

View the created table
```{r}
print(days_sleep_day)
```

Combining days_daily_activity and days_sleep_day on 'day_name'
```{r}
combined_activity_sleep <- left_join(days_daily_activity, days_sleep_day, 'day_name')
```

View the combined table
```{r}
print(combined_activity_sleep)
```

Changing the format of date - weightlog_info
```{r}
weightlog_info$Date = mdy_hms(weightlog_info$Date)
```

Adding a new column 'day_name' for the corresponding dates
```{r}
weightlog_info <- weightlog_info %>%
  mutate(day_name = wday(Date, label = TRUE))
```

## Analysing the Data -

Creating a summarized daily_activity table by Id
```{r}
id_daily_activity <- daily_activity %>%
  group_by(Id) %>%
  summarise(no_days = max(ActivityDate) - min(ActivityDate), total_steps = sum(TotalSteps), total_distance = sum(TotalDistance), tracker_dis = sum(TrackerDistance),
            log_activities_dist = sum(LoggedActivitiesDistance), Very_active_dist = sum(VeryActiveDistance), mod_active_dist = sum(ModeratelyActiveDistance), 
            light_active_dist = sum(LightActiveDistance), sed_active_dist = sum(SedentaryActiveDistance), very_active_min = sum(VeryActiveMinutes), 
            fairly_active_min = sum(FairlyActiveMinutes), lightly_active_mins = sum(LightlyActiveMinutes), sed_min = sum(SedentaryMinutes), 
            total_calories = sum(Calories)) %>% 
  select(Id, no_days, total_steps, total_distance, tracker_dis, log_activities_dist, Very_active_dist, mod_active_dist, light_active_dist, sed_active_dist, 
         very_active_min, fairly_active_min, lightly_active_mins, sed_min, total_calories) 
```

Creating a summarized sleep_day table by Id
```{r}
id_sleep_day <- sleep_day %>%
  group_by(Id) %>%
  summarise(no_days = (max(SleepDay) - min(SleepDay)) / 86400,  total_sleep_rec = sum(TotalSleepRecords), total_min_asleep = sum(TotalMinutesAsleep), 
            total_time_bed = sum(TotalTimeInBed)) %>% 
  select(Id, no_days, total_sleep_rec, total_min_asleep, total_time_bed)
```

Creating a summarized weightlog_info table by Id
```{r}
id_weightlog_info <- weightlog_info %>%
  group_by(Id) %>%
  summarise(no_days = (max(Date) - min(Date)) / 86400, max_weight = max(WeightKg), min_weight = min(WeightKg), max_fat = max(Fat), min_fat = min(Fat),
            max_bmi = max(BMI), min_bmi = min(BMI)) %>%
  select(Id, no_days, max_weight, min_weight, max_fat, min_fat, max_bmi, min_bmi)

```

View tables created by Id
```{r}
View(id_daily_activity)
View(id_sleep_day)
View(id_weightlog_info)
```

## Visualizing the Data -

total_steps Vs. day_of_week
```{r}
ggplot(data = combined_activity_sleep, aes(x = day_name, y = total_steps, fill = day_name)) + geom_bar(stat = "identity")
```

Observation: Calories has a positive correlation with no. of steps

day_of_week Vs. mins_asleep
```{r}
ggplot(data = combined_activity_sleep, aes(x = day_name, y = tot_min_asleep , fill = day_name)) + geom_bar(stat = "identity")
```

Observation: Total mins asleep also has a positive correlation with no. of steps

total_calories Vs. day_of_week
```{r}
ggplot(data = combined_activity_sleep, aes(x = day_name, y = total_calories, fill = day_name)) + geom_bar(stat = "identity")
```

Observation: Calories also has a positive correlation with no. of steps

total_steps Vs. sedentary_mins
```{r}
ggplot(data = daily_activity) + geom_smooth(mapping = aes(x = TotalSteps, y = SedentaryMinutes))
```

 geom_smooth()` using method = 'loess' and formula 'y ~ x'
Observation: Sedentary minutes reduces as total steps increases, the least sedentary minutes are b/w 10,000 to 20,000 steps & spikes up after 20,000 steps

total_mins_asleep Vs. total_time_in_bed
```{r}
ggplot(data = sleep_day) + geom_smooth(mapping = aes(x = TotalMinutesAsleep, y = TotalTimeInBed))

#### geom_smooth()` using method = 'loess' and formula 'y ~ x'
```
Observation: Total time in bed increases with total mins asleep

## Conclusions:
- Users that used the smart devices to track steps were found to have their highest record on Tuesday followed by Wednesday.
- Time asleep increases with the increase in steps.
- The tendency to use smart device to track weight information is very low in the users of such devices.
- The brand specialises in Women's products and aims at improving their health by providing the facility to track reproductive health.

## Recommendations:
- Bellabeat should form a day-wise weekly social media strategy for customers and should encourage the customers to use the health bands for step tracking for other days too.  
- Bellabeat can motivate it's users for tracking more of their sleep patterns on the app by sending reminders each night to improve them. 
- Bellabeat should encourage the customers to use their app to track their food habits and behaviours, give them advice on healthy eating and excercising to control their calory intake and burn. This will motivate customers to use their devices even more.
- Bellabeat should interact to customers via posting health awareness content on their social media, and explain how food, sleep, excercise and body weight can influence on hormones and all over their health. This can create a new customer base for the company.  

